---
title: "Settings_BoxModel_OceanMargin"
author: "Anne Kruijt"
date: "25 januari 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```
This code was developed by Karline Soetaert and Anne Kruijt, to simulate the transfer of carbon and potentially other chemical tracers along the coastal shelf, represented by a flexible number of boxes. 

The original code is split into several individual scripts, each for a distinct step within the set-up and execution of the code.
THIS script contains the set-up of the model domain and the boundary conditions.


## Note to self: Ik moet goed gaan nadenken over het resolutie vraagstuk: als ik nx omhoog doe gaat het fout bij de calcificatie in het model. Waarom? is die niet goed gekoppeld aan daadwerkelijke hoeveelheid beschikbare bouwstenen? Raken de 'volumes' van de boxen vol? Ik moet hier met Karline over praten, maar zal iig:
- de resolutie omlaag doen van 10 naar 5, kijken wat dat doet: GEEN ERROR
- resolutie met 1 of 2 boxen omhoog (let op, de verdeling in bodembox-dieptes moet ook aangepast): WEL EEN ERROR
- rates naar p/ms doen ipv /s, kijken hoe dat de boel beinvloedt: WEL EEN ERROR
- diepte surface box aanpassen van 5 naar 10
- resolutie hetzelfde en area omhoog: GEEN ERROR (bij verdubbeling area)

##Packages used
```{r, message = FALSE}
install.packages("docstring")
library(docstring)
require(rootSolve)
require(deSolve)

list.of.packages <- c("seacarb", "ggplot2", "grid")

new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)
invisible(lapply(list.of.packages, library, character.only = TRUE))
rm(list.of.packages,new.packages)
```

## Senario model settings

Here the dimensions, bathymetry and environmental conditions (temperature, salinity, light) of the modelled ocean margin can be set. 

To simulate carbon transfer across the global oceanic margin we attribute modern day estimates of global ocean margin dimensions. 

The ocean margin bathymetry and environmental settings are prescribed separately, so the user needs to decide what a realistic combination between box depth, area and enviromental parameters would be, and be aware of the effect of choices made below.


```{r settings }

##---Overal Settings: Global ocean margin ---#
Area = 5.00e6 # [km2]
Qriver =  7.41e3 # [km3/s]
Eh =  0  # horizontal dispersion [km3/s]

##--- Bathymetry ---#
# Bathymetry is created by assuming three (increasing) depth ranges for the bottom boxes, each depth range with a distinct environment. Suggested settings for different environments can be found in separate excel file ("NAME").
d_surf = 5  # [m]
d_bottom_s = 10  # where corals grow
d_bottom_m = 50
d_bottom_d = 140

# Setting fraction of ocean margin for each depth range
bs <- 1/10 #1/5 #bottom shallow
bm <- 4/10 #2/5 #bottom middle
bd <-  5/10 #2/5 #bottom deep

##--- Temperature settings ---#
Ts = 26 # surface box temperature [deg Celcius]
Tbs = 25 # bottom shallow
Tbm = 20 # bottom middle
Tbd = 15 # bottom deep

##--- Light settings ---#
E0 = 10 # surface light, [mol quanta/m2/day]
kpar = 0.3 # or 0.5 for coastal waters? 


Envparms <- c(Ts, Tbs, Tbm, Tbd, E0, kpar)

Box_settings <- function(nx, p){
  #' @title Box settings
  #' @description This function assigns the appropriate dimensions and environmental settings to each box of the model. These dimensions and environmental conditions have been given by the user.
  #' @param x The number of boxes in the horizontal (nx)
  #' @param y A list containing the environmental parameters to be assigned
  #' @return A list with the items: temperature, light availability, volume, salinity, and depth of each box. Each item is itself a vector of lenght 2 * nx, containing the value of this item for each individual box

  # Salinity
  S = cbind(c(5, seq(from=25, to=35, length.out=nx + 1)), seq(from=30, to=35, length.out=nx +2)) # salinity matrix, first column representing surface boxes, second column bottom boxes

  # Temperature
  with(as.list(p),{
  Temp_surf <- c(rep(Ts, times =nx))
  Temp_bottom <- c(rep(Tbs, times =nx*bs), rep(Tbm, times = nx*bm), rep(Tbd, times = nx*bd))
  Temp <- c(Temp_surf, Temp_bottom)
  
  #Light, average over box depth    
  z <- seq(0, (d_bottom_d+d_surf+10), by=1)   # depth in meters; make sure the z range is at least as large as the maximum depth of the model environment
  #   for (i in z ){
  # E <- E0 * exp(-kpar *z)
  # E_s <- c(rep(mean(E[0: (d_surf)]), times =nx)) 
  # E_bs <- c(rep(mean(E[((d_surf+1): (d_bottom_s+d_surf))]), times =nx/5))
  # E_bm <- c(rep(mean(E[((d_surf+d_bottom_s+1): (d_bottom_m+d_surf))]), times =2*(nx/5)))
  # E_bd <- c(rep(mean(E[((d_surf+d_bottom_m+1): (d_bottom_d+d_surf))]), times =2*(nx/5)))
  # Ebox <-c(E_s, E_bs, E_bm, E_bd) # mol quanta/m2/s
  #   }
  
  # Light that reaches bottom of box
  for (i in z ){
  E <- E0 * exp(-kpar *z)
  E_s <- c(rep(E[d_surf], times =nx)) 
  E_bs <- c(rep(E[(d_bottom_s+d_surf)], times =nx*bs))
  E_bm <- c(rep(E[(d_bottom_m+d_surf)], times =nx*bm))
  E_bd <- c(rep(E[(d_bottom_d+d_surf)], times =nx*bd))
  Ebox <-c(E_s, E_bs, E_bm, E_bd) # [mol quanta/m2/s]
  }
  
  # Depths in m
  Depth_b <- c(rep(d_bottom_s, times = nx*bs), rep(d_bottom_m, times = nx*bm),rep(d_bottom_d, times = nx*bd) )
  Depths <- cbind(d_surf, Depth_b)
  
  # Volume 
  # recalculate to km3, this unit is needed for the calculation of flows in hydrodynamics-script
  Vsurf <- c(rep(d_surf/1000*(Area/nx)))
  Vbottom <-c(rep(d_bottom_s/1000*(Area/nx), times = nx*bs), rep(d_bottom_m/1000*(Area/nx), times = nx*bm),rep(d_bottom_d/1000*(Area/nx), times = nx*bd) )
  BoxVolume <- cbind(Vsurf, Vbottom)
  
  list('Temp'=Temp, 'Ebox' = Ebox, 'BoxVolume'= BoxVolume, 'S'= S, "Depths" = Depths)
  
  })
}


```


## Determining appropriate river DIC and TA, based, knowing pH should be 7.6 and salinity of water low (e.g. 5)
```{r Boundary codition determination }

RiverTA = 895
RiverDIC = 910
carb_solverriv<-carb(flag =15, (RiverTA/1000000), (RiverDIC/1000000), S=5)
pHriv <- as.numeric(carb_solverriv["pH"])

```

```{r calculating and visualising the box-conditions}

## Calculating boxconditions
nx =  10
Boxconditions <-Box_settings(nx,Envparms)
Temp <- Boxconditions[["Temp"]]  # in deg C
E <- Boxconditions[["Ebox"]]  # in mol quanta/m2/s
BoxV <- Boxconditions[["BoxVolume"]] # in km3
S <- Boxconditions[["S"]]

## checking all box conditions
Bathymetry <- c(rep(d_bottom_s, times = nx*bs), rep(d_bottom_m, times = (nx*bm)),rep(d_bottom_d, times = (nx*bd)) )*(-1)  # in meters
x <-seq(1,nx)
plot(x, Bathymetry, type = 'l', main= "Ocean margin bathymetry", xlab = "Box number", ylab= "Depth")
plot(x, Temp[1:nx], type = "l", main = "Box temperatures", ylab ="Temperature", col="red")
lines(x,Temp[(nx+1):(2*nx)],col="green")
plot(x, S[2:(nx+1), 1], type ="l", main = "Box salinities", ylab = "Salinity", col = "red")
lines(x, S[2:(nx+1), 2],col="green")


# Saving plots to a folder on Git


jpeg(file= "../results/figures/OceanMarginBathym.jpeg")
  plot(x, Bathymetry, type = 'l', main= "Ocean margin bathymetry", xlab = "Box number", ylab= "Depth")
dev.off()

```


```{r Temperature scenarios}

# function to determine the volume fraction of each depth environment
VolumeFractionFun <- function(BoxV){
  TotalV <- sum(BoxV)
  fractionSurf <- sum(BoxV[,1])/TotalV
  fractionBs <- sum(BoxV[1:(nx*bs),2])/TotalV
  fractionBm <- sum(BoxV[(nx*bs+1): (nx*bm+ nx*bs),2])/TotalV
  fractionBd <- sum(BoxV[(nx*bs+nx*bm+1):(nx),2])/TotalV
  list('Surface fraction'=fractionSurf, 'Shallow bottom fraction' = fractionBs, 'Middle bottom fraction'= fractionBm, 'Deep bottom fraction'= fractionBd)
}

fracs <-VolumeFractionFun(BoxV)

Tsmax <- 26
Tbsmax <-25
Tbmmax <-20
Tbdmax <-15

AvTemp <- Tsmax*fracs$`Surface fraction` + Tbsmax* fracs$`Shallow bottom fraction` + Tbmmax*fracs$`Middle bottom fraction` + Tbdmax*fracs$`Deep bottom fraction`
AvTemp_min3 <- (Tsmax-3)*fracs$`Surface fraction` + (Tbsmax-3)* fracs$`Shallow bottom fraction` + (Tbmmax-3)*fracs$`Middle bottom fraction` + (Tbdmax-3)*fracs$`Deep bottom fraction`
AvTemp_min7 <-(Tsmax-7)*fracs$`Surface fraction` + (Tbsmax-7)* fracs$`Shallow bottom fraction` + (Tbmmax-7)*fracs$`Middle bottom fraction` + (Tbdmax-7)*fracs$`Deep bottom fraction`


sec_to_year <- 60 * 60 * 24 * 365

```